---
title: "bar_processing_script"
output: html_document
date: "2023-07-06"
---

```{r setup, include=FALSE}

# Library load ==== 

libraries <- c("tidyverse", "janitor", "data.table", "readxl")
lapply(libraries, require, character.only = T) # the following line of code reads in the required libraries for this script

```

```{r}
# VARIABLES TO CHANGE

data_directory <- r"(C:/Users/u453623/Documents/OFFLINE/bar_template_processing/data/)" # THIS LINE OF CODE POINTS TO WHERE ALL THE DATA WILL BE STORED, **CHANGE AS NEEDED** 

#TODO: maybe utilise relational data pathing so that this would work wherever you paste the project? Unsure how this would be helpful when other people might want to use the script
```

#Processing of BAR data - Script explained

This script was written for the purpose of collating multiple BAR templates into a single dataset

This involves tgree steps:

 1) Reading in all the excel spreadsheets in one place
 2) Formatting the excel spreadsheet into a consistent format
 3) Joining these formatted entries into a single data set

Throughout this script, comments are used to explain what each function does, however (hopefully) all that would need to be done is to run the script and it should work.

## 1) Reading in the spreadsheets

```{r}

reading_spreadsheets <- function(data_directory) { 
  names <- list.files(data_directory)

all_files <- paste0(data_directory, names) # This line of code saves the names of all the files from the folder specified above into a variable called "all_files"

df.list <- lapply(all_files, read_excel, sheet = 2) # Apply function read_xlsx to all files in file directory. Stored in a list . We are specifically only interested in the 2nd sheet, which as all the data 

return(df.list)

}

df.list <- reading_spreadsheets(data_directory)

```

## 2) Formatting the spreadsheets 

We have several functions that we neeed to set up. Each chunk will correspond to one of these functions. 

### General formatting 

```{r}
# We want to move the first row of entries to be the column names, we could also do with tidying up their names into a consistent format. The janitor package is great for these tasks.



# This function takes a given list, and applies the functions row_to_names and clean_names to move the column names from row 1, and cleanes them in a consistent format. 

df.list <- df.list |> 
    lapply(janitor::row_to_names, 1) |> 
    lapply(janitor::clean_names)
  
```

### SIC codes function

```{r}

# there is extra datafields in the sic_codes column, explaining how to look up a sic code for those completing the template. We do not need that for the purpose of analysing the data, so it will be removed.  

remove_SIC_codes_info <- function(data) {
  data %>%
    mutate(x11_sic_code = ifelse(row_number() > 1, NA, x11_sic_code))
}

df.list <- df.list |> 
  lapply(remove_SIC_codes_info)

```

### Pivot Wider

```{r}

# First, we define a vector of variables that need to flattned. If you have a look at the template, these vars are the ones that have multiple questions in the same column. The questions are stored in 'names_from_vars' and the corresponding answers will be in 'values_from_vars'

# after we flatten this, every column will have an answer for only one question

names_from_vars <- c("x3_b_location_if_uprn_is_missing", "x4_full_property_address", "x5_assessor_details", "x13_property_occupancy", "x17_heat_demand_m_wh_year", "na_6", "x20_heat_network_connection", "na_8", "x26_expected_impact_of_changes")
values_from_vars <- c("na", "na_2", "na_3", "na_4", "na_5", "x18_a_primary_fuel_demand_unit_year", "na_7", "x22_a_primary_heating_technology", "na_9")
```


```{r}

perform_pivot_wider <- function(data, key_vars, value_vars) {

    df <- data
  
  for (i in seq_along(key_vars)) {
    
    pivot_key <- key_vars[i]
    pivot_value <- value_vars[i]
    
    #print(c(pivot_value, pivot_key))
    
    df <- df %>%
      pivot_wider(names_from = pivot_key, 
                  values_from = pivot_value, 
                  names_repair = "unique")
  }
  
  return(df)
}

df.list <- df.list |> 
  lapply(perform_pivot_wider, key_vars = names_from_vars, value_vars = values_from_vars)

```

### Compressing final outcome

```{r}

compress_data_frame <- function(df) {
  # Extract non-NA values
  non_na_values <- na.omit(unlist(df))
  
  # Create a new data frame with a single row
  compressed_df <- data.frame(t(non_na_values))
  
  return(compressed_df)
}

df.list <- df.list |> 
  lapply(compress_data_frame)
```

# Examine the data.frames in each list

```{r}

df1 <- df.list[[1]]
df2 <- df.list[[2]]

df1 %>% select(order(colnames(df1)))
```


### Order Columns 


```{r}

```

